---
title: "Passing Data Between R and Python"
author: "Priyanka Sodhi"
date: 2026-09-27
---

## Introduction

This post demonstrates how R and Python can be used together in the same Quarto document. I create an object in R, use Python to work with that object, and then bring the Python result back into R.

The dataset used is the built-in `iris` dataset in R.

Data source: [R iris dataset](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/iris.html).

## Creating data in R

First, I create a vector in R containing the petal lengths for the iris dataset.

```{r}
#| message: false
#| warning: false

data(iris)
library(reticulate)


petal_lengths <- iris$Petal.Length

head(petal_lengths)
```

The `petal_lengths` object is created in R and contains the petal length measurements for all 150 observations.

## Using the R object in Python

The Python chunk can access the R object using `r.petal_lengths`.

```{python}
import numpy as np

python_petal_lengths = np.array(r.petal_lengths)

python_mean = python_petal_lengths.mean()

python_mean
```

Python receives the R vector and calculates its mean petal length.

## Passing the Python result back to R

The Python result can then be accessed from R using `py$python_mean`.

```{r}
python_mean_from_r <- py$python_mean

python_mean_from_r
```

The result calculated in Python is now available in R.

## Visualizing the data

Finally, I use the original R dataset to visualize the petal length measurements across species.

```{r}
#| label: fig-r-python-petal-length
#| fig-cap: "Petal length measurements across iris species."

boxplot(
  Petal.Length ~ Species,
  data = iris,
  xlab = "Iris species",
  ylab = "Petal length",
  main = "Petal Length by Species"
)
```

The R and Python chunks in this post share objects through `reticulate`, allowing both languages to be used together in the same computational document.
